---
title: "EssentialTnSeq"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    theme: flatly
    highlight: kate
vignette: >
  %\VignetteIndexEntry{EssentialTnSeq}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  comment = "",
  collapse = TRUE
)
```

<hr>

## Installation and Setup
The `devtools` package is required for installation of EssentialTnSeq.
Additionally, the functions included with this package are designed to work with
(and in fact make use of) the `tidyverse` suite of packages/functions.
```{r setup}
# First, install dependencies devtools and tidyverse
# install.packages(c("devtools", "tidyverse"))

# Then install and load EssentialTnSeq
# devtools::install_github("travis-m-blimkie/EssentialTnSeq")
# Also load the tidyverse
suppressPackageStartupMessages({
  library(EssentialTnSeq)
  library(tidyverse)
})
```


## Introduction
This package was designed to facilitate the analysis of TnSeq data;
specifically, the results of Tradis or Transit/Gumbel essentiality pipelines.
The package includes three functions which are designed to be used together to
create a complete workflow, from Tradis or Transit/Gumbel outputs to lists of
essential genes.  
The three functions and examples of their usage are demonstrated in this vignette.  

## Tradis
### 0. Input Data
As mentioned previously, this package and its functions are built to make use of the output from essenitality analysis of TnSeq data from Tradis or Gumbel.  
Input files for the Tradis side of this essentiality pipeline should have been
generated as a result of calling `tradis_gene_insert_sites` from the Tradis
suite of functions. Files should be named something like
"condition1_genomeRef.tradis_gene_insert_sites.csv.all.csv". Note final element
of the file name ("csv.all.csv") is used by the EssentialTnSeq import function
to identify relevant files, and as such should be consistent with what is
denoted here.  
An example directory containing Tradis results is included below. 

[]()


### 1. Reading Data
This first step is covered by the function `et_readFiles()`. The required
arguments for this function are detailed below, along with an example (keeping
with the directory shown above).  
Arguments:

* tool - one of "Tradis" or "Gumbel". For this current workflow, we will use "Tradis".
* conditions - a list or character vector of condition names. These should match
file names, and not overlap with one another. For example, if you have bacteria
grown in different media, along with treatment/no treatment, they should be
named like: "media1_untreated" and "media1_treated". In this case, using just
"media1" in place of "media1_untreated" means the incorrect number of files
would be identified (as "media1" also matches "media1_treated".)
* reps - the number of replicates for the experimental conditions. Should be a
character vector, like `c("r1", "r2", "r3")`. It should also match the number of
files for each condition, otherwise file input will not function properly (you
will get an error).
* data_folder - the directory containing all of the data files to be used (i.e.
all conditions and replicates).

The code below utilizes the preview data shown above, reading in the files with
the `et_readFiles()` function and previously described arguments.

```{r et_readFiles, eval=FALSE}
my_conditions <- c("media1_untreated", "media1_treated")
my_reps <- c("r1", "r2", "r3")

raw_data <- et_readFiles(tool = "Tradis", 
                         conditions = my_conditions, 
                         reps = my_reps, 
                         data_folder = "Data_dir")
```




